### 2022-10-11 Papers/ICCC-2023/sections/design.tex
stuff

An important consideration with implementing a CTFE First compiler, is what limitations to put on modifications of user code. %todo: longer
Depending on language design, it may be possible to introduce circular references between the functions that modify the codebase.
For example if function \lstinline{A} is modified by function \lstinline{B} and \lstinline{B} invokes \lstinline{A}, the behaviour of the program is unpredictable.
This problem will only be magnified by larger program sizes.

One of the possible solutions to the above mentioned problem is to limit which functions can be invoked at compiletime.
C-=-1 allows code within a compiletime context to invoke procedures only from other packages, declared explicitly as dependencies.
It additionally prohibits modification of dependencies.
Therefore its is impossible for a function to modify a procedure it depends on.% todo: wording, example

>>>>

\begin{figure}
	\includegraphics[width=6cm]{pictures/circular-function-reference.jpg}
	\caption{Example of circular reference in meta code}
	\label{circular-function-reference}
\end{figure}

An important consideration with implementing a CTFE First compiler, is what code can be executed at compile-time. %todo: longer
Depending on language design, it may be possible to introduce circular references between the functions that modify the codebase.
Figure \ref{circular-function-reference} contains a diagram of such a scenario.
\lstinline{A}, \lstinline{B} and \lstinline{C} are all functions.
If function \lstinline{B} is modified by function \lstinline{C} and \lstinline{C} invokes \lstinline{B}, the behaviour of function \lstinline{A} is unpredictable.
This problem will only be magnified by larger program sizes.

One of the possible solutions to the above mentioned problem is to limit which functions can be invoked at compiletime.
C-=-1 allows code within a compiletime context to invoke procedures only from other packages, declared explicitly as dependencies.
Circular references between packages, as in most other languages, are forbiden.
C-=-1 additionally prohibits modification of dependencies.
Therefore its is impossible for a function to modify a procedure it depends on.% todo: wording, example

### 2022-10-11 Papers/ICCC-2023/sections/implementation.tex
stuff

Languages that seek to exploit CTFE First approach in design of their compilers, should define a set of data structures to describe user code, i.e. an Semantic Model.
It will allow the programmer to interact and mainpulate the structure of the program at compile-time.
The Semantic model designed and implemented for C-=-1 is described in chapter \ref{semantic_model}.

@@

To deal with this complexity, C-=-1 semantic model for functions is build around two interfaces: \lstinline{IInstruction} and \lstinline{IExpression} and their concrete implementations.

@@

Certain other intrinsic operations were defined using external dependencies.
C-=-1 memory managment library, in the runtime context, builds on a simple interface capable of allocating and deleting a continous buffer.
It is comprised of two functions: \lstinline{unsafe_new} and \lstinline{delete}.
In the standard library, they are explicitly mapped to \lstinline{malloc} and \lstinline{free} functions from the C runtime.

>>>>

Languages that seek to exploit CTFE First approach in design of their compilers, should define a set of data structures to describe user code, i.e. a Semantic Model. %todo: languages are not people
It will allow the programmer to interact and mainpulate the structure of the program at compile-time.
The Semantic model designed and implemented for C-=-1 is described in chapter \ref{semantic_model}.

@@

To deal with this complexity, C-=-1 semantic model for functions is build around two interfaces: \lstinline{IInstruction} and \lstinline{IExpression} and their concrete implementations.
Every category of instruction or expression is represented by its own type.
The user may then analyze the structure of the program, using a dynamic type conversion mechanism similar to C++ \lstinline{dynamic_cast} \cite{ISO:cpp98}.

To aid in generating meaningfull error messages, all elements of the semantic model, have a \lstinline{sourcePointers}.
It is a simple structure, that contains the filename and the line number of the expression or instruction.
This information can be passed to compiler intrinsic functions, such as \lstinline{raiseError}, to generate error messages for the user.
Listing \ref{lst:noDiscardCm1} contains a example use of this functionality.

Component responsible for creating the semantic model are a major part of the compiler.
There are two operations that this module performs: building the definition of the types used to describe a program, and creating an instance of the model, given semantic information.
C-=-1 compiler has a hard-coded definition of its base library.
It contains definitions of primitive types and types used to build the semantic model of a program.
The description of this library must be built manually, as it is very closley integrated with the compiler.

@@

Certain other intrinsic operations were defined using external dependencies.
C-=-1 memory managment library, in the runtime context, uses a simple interface capable of allocating and deleting a continous buffer.
It is comprised of two functions: \lstinline{unsafe_new} and \lstinline{delete}.
In the standard library, they are explicitly mapped to \lstinline{malloc} and \lstinline{free} functions from the C runtime.
